---
title: "human vs algorithm data processing"
author: "LGCarlson"
date: "10/9/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(here)
library(zoo) 

check = function(x){summary(as.factor(x))}

#set personalized plotting theme
LGCtheme <- theme_light() + theme(panel.background = element_blank(), 
                panel.grid = element_blank(), 
                axis.text = element_text(size = 13), 
                axis.title = element_text(size = 14), 
                legend.text = element_text(size = 13), 
                legend.title = element_text(size = 14), 
                strip.background = element_blank(), 
                strip.text = element_text(size = 13, color = "black"))

theme_set(theme_light())
knitr::opts_chunk$set(echo = TRUE)
```

Read in the small amount of Carlin data that Brandon did mark seasonal max/mins for

```{r,message=F}
homewater_human<-read_csv(here::here("raw_data", "homewater_humanmarked.csv"))
```

Create a populate a new dataframe with increments and circuli # for seasonal increments based on human seasonal markers

```{r}
human_marked<-data.frame(matrix(NaN,nrow=nrow(homewater_human),ncol=35))

colnames(human_marked)<-c("JoinID","ReleaseYear","markersource","M1.loc","M2.loc","FMC","FSM","FWM","SSM","SWM",
                    "FMC.circ","smolt.circ","marine.circ","total.circ","FSM.circ","FWM.circ","SSM.circ","SWM.circ",
                    "FS.circ","FW.circ","PS.circ","SS.circ","SW.circ","Y2.circ","M1","M2",
                    "smolt.incr","marine.incr","total.incr","FS.incr","FW.incr",
                    "PS.incr","SS.incr","SW.incr","Y2.incr")

# loop to calculate growth increment data
for(i in 1:nrow(homewater_human)){
  
   # Select row and locate column numbers (13:154) for circuli data (may need to change this for different data sets)
  vect<-as.numeric(homewater_human[i,15:140])
  
  # Get rid of NA at the end of the vector
  vect<-vect[!is.na(vect)]
  
  # Intercirculi spacings
  spacings<-diff(vect)
  
  # ID
  human_marked$JoinID[i]<-homewater_human$JoinID[i]
  
  # year
  human_marked$ReleaseYear[i]<-homewater_human$ReleaseYear[i]
  
  # markersource
  human_marked$markersource[i]<-"human"
  
  # M1
  human_marked$M1.loc[i]<-homewater_human$M1[i]
  
  # M2
  human_marked$M2.loc[i]<-homewater_human$M2[i]
  
  # FMC
  human_marked$FMC[i]<-homewater_human$FMC[i]

  
  
    ####################################################################################
  
  # Depending on which growth markers are present, certain growth increments will be identifiable
  
  # is FMC present? if not, keep this row NaN and then go to next row
  if(is.na(homewater_human$FMC[i])){
    next
  }
  
  ####################################################################################
  
  # if FMC is present, calculate these growth markers, increments and circuli
  
  # First marine circuli (FMC) circuli number
  FMC<-(which(homewater_human$FMC[i]<=vect)[1])-1

  human_marked$FMC.circ[i]<-FMC 
  
  # smolt increment (focus to last circuli before FMC)
  human_marked$smolt.incr[i]<-sum(spacings[1:(FMC-1)])
  
  # marine growth (FMC to end)
  human_marked$marine.incr[i]<-sum(spacings[FMC:length(spacings)])
  
  # total growth (entire scale from 1st circuli to the end)
  human_marked$total.incr[i]<-sum(spacings)
  
  # number of smolt circuli
  human_marked$smolt.circ[i]<-length(spacings[1:(FMC-1)])
  
  # number of marine circuli
  human_marked$marine.circ[i]<-length(spacings[FMC:length(spacings)])
  
  # number of total circuli
  human_marked$total.circ[i]<-length(spacings)
  
  ####################################################################################
  
  # Depending on which growth markers are present, certain growth increments will be identifiable
  
  # is M1 present? if not, go to next row
  if(is.na(homewater_human$M1[i])){
    next
  }
  
  ####################################################################################
  
  # if M1 is present, calculate these increments and circuli
  
  # M1 circuli number
  M1<-(which(homewater_human$M1[i]<=vect)[1])-1
  human_marked$M1[i]<-M1
  
  FSM<-(which(homewater_human$SM1[i]<=vect)[1])-1
  
  # First summer maximum (FSM) circulus number
  human_marked$FSM.circ[i]<-FSM
  human_marked$FSM[i]<-homewater_human$SM1[i]
  
  # First summer at sea growth (FMC to FSM)
  human_marked$FS.incr[i]<-sum(spacings[FMC:FSM])
  
  # First winter at sea growth (first circuli after FSM to M1)
  human_marked$FW.incr[i]<-sum(spacings[(FSM+1):M1])
  
  # First winter minimum (FWM)
  FWM<-(which(homewater_human$WM1[i]<=vect)[1])-1
  
  # First winter minimum (FWM) circulus number
  human_marked$FWM.circ[i]<-FWM
  human_marked$FWM[i]<-homewater_human$WM1[i]
  
  # Post-smolt growth (FMC to M1)
  human_marked$PS.incr[i]<-sum(spacings[FMC:M1])
  
  # First summer, number of circuli
  human_marked$FS.circ[i]<-length(spacings[FMC:FSM])
  
  # First winter, number of circuli
  human_marked$FW.circ[i]<-length(spacings[(FSM+1):M1])
  
  # Post-smolt, number of circuli
  human_marked$PS.circ[i]<-length(spacings[FMC:M1])
  
  
  ####################################################################################
  
  # Depending on which growth markers are present, certain growth increments will be identifiable
  
  # is M2 present? if not, go to next row
  if(is.na(homewater_human$M2[i])){
    next
  }
  
  ####################################################################################
  
  # if M2 is present, calculate these increments and circuli
  
  # M2 circuli number if present
  M2<-(which(homewater_human$M2[i]<=vect)[1])-1
  human_marked$M2[i]<-M2
  
  # Second summer maximum
  SSM<-(which(homewater_human$SM2[i]<=vect)[1])-1
  
  # Second summer maximum (SSM) circulus number
  human_marked$SSM.circ[i]<-SSM
  human_marked$SSM[i]<-homewater_human$SM2[i]
  
  # Second summer at sea growth (first circuli after M1 to SSM)
  human_marked$SS.incr[i]<-sum(spacings[(M1+1):SSM])
  
  # Second winter at sea growth (first circuli after SSM to M2)
  human_marked$SW.incr[i]<-sum(spacings[(SSM+1):M2])
  
  
  
  # Second winter minimum (SWM)
  SWM<-(which(homewater_human$WM2[i]<=vect)[1])-1
  
  # Second winter minimum (SWM) circulus number
  human_marked$SWM.circ[i]<-SWM
  human_marked$SWM[i]<-homewater_human$WM2[i]
  
  
  # Year 2 growth (first circuli after M1 to M2)
  human_marked$Y2.incr[i]<-sum(spacings[(M1+1):M2])
  
  # Second summer, number of circuli
  human_marked$SS.circ[i]<-length(spacings[(M1+1):SSM])
  
  # Second winter, number of circuli
  human_marked$SW.circ[i]<-length(spacings[(SSM+1):M2])
  
  # Year-two number of circuli
  human_marked$Y2.circ[i]<-length(spacings[(M1+1):M2])
  
}
```



Create a populate a new dataframe with increments and circuli # for seasonal increments as calculated by the algorithm

```{r}
alg_marked<-data.frame(matrix(NaN,nrow=nrow(homewater_human),ncol=35))

colnames(alg_marked)<-c("JoinID","ReleaseYear","markersource","M1.loc","M2.loc","FMC","FSM","FWM","SSM","SWM",
                    "FMC.circ","smolt.circ","marine.circ","total.circ","FSM.circ","FWM.circ","SSM.circ","SWM.circ",
                    "FS.circ","FW.circ","PS.circ","SS.circ","SW.circ","Y2.circ","M1","M2",
                    "smolt.incr","marine.incr","total.incr","FS.incr","FW.incr",
                    "PS.incr","SS.incr","SW.incr","Y2.incr")

# loop to calculate growth increment data
for(i in 1:nrow(homewater_human)){
  
  # Select row and locate column numbers (13:154) for circuli data (may need to change this for different data sets)
  vect<-as.numeric(homewater_human[i,15:140])
  
  # Get rid of NA at the end of the vector
  vect<-vect[!is.na(vect)]
  
  # Smoothed circuli spacing, 5 pairs
  vectavg<-zoo::rollapply(data=vect,width=5,FUN=mean,partial=T,align="center",na.rm=T)
  
  # Intercirculi spacings
  spacingavg<-diff(vectavg)
  
  # ID
  alg_marked$JoinID[i]<-homewater_human$JoinID[i]
  
  # year
  alg_marked$ReleaseYear[i]<-homewater_human$ReleaseYear[i]
  
  # markersource
  alg_marked$markersource[i]<-"algorithm"
  
  # M1
  alg_marked$M1.loc[i]<-homewater_human$M1[i]
  
  # M2
  alg_marked$M2.loc[i]<-homewater_human$M2[i]
  
  # FMC
  alg_marked$FMC[i]<-homewater_human$FMC[i]
  
  
    ####################################################################################
  
  # Depending on which growth markers are present, certain growth increments will be identifiable
  
  # is FMC present? if not, keep this row NaN and then go to next row
  if(is.na(homewater_human$FMC[i])){
    next
  }
  
  ####################################################################################
  
  # if FMC is present, calculate these growth markers, increments and circuli
  
  # First marine circuli (FMC) circuli number
  FMC<-(which(homewater_human$FMC[i]<=vect)[1])-1

  alg_marked$FMC.circ[i]<-FMC 
  
  # smolt increment (focus to last circuli before FMC)
  alg_marked$smolt.incr[i]<-sum(spacingavg[1:(FMC-1)])
  
  # marine growth (FMC to end)
  alg_marked$marine.incr[i]<-sum(spacingavg[FMC:length(spacingavg)])
  
  # total growth (entire scale from 1st circuli to the end)
  alg_marked$total.incr[i]<-sum(spacingavg)
  
  # number of smolt circuli
  alg_marked$smolt.circ[i]<-length(spacingavg[1:(FMC-1)])
  
  # number of marine circuli
  alg_marked$marine.circ[i]<-length(spacingavg[FMC:length(spacingavg)])
  
  # number of total circuli
  alg_marked$total.circ[i]<-length(spacingavg)
  
  ####################################################################################
  
  # Depending on which growth markers are present, certain growth increments will be identifiable
  
  # is M1 present? if not, go to next row
  if(is.na(homewater_human$M1[i])){
    next
  }
  
  ####################################################################################
  
  # if M1 is present, calculate these growth markers, increments and circuli
  
  # M1 circuli number
  M1<-(which(homewater_human$M1[i]<=vect)[1])-1
  alg_marked$M1[i]<-M1
  
  # First summer maximum (FSM) circuli number
  # (subtract 1 from FMC because it is already included in indexing)
  FSM<-(FMC-1)+which(spacingavg[FMC:M1]==max(spacingavg[FMC:M1]))[1]
  
  # First summer maximum (FSM) circulus number
  alg_marked$FSM.circ[i]<-FSM
  alg_marked$FSM[i]<-vect[FSM]
  
  # First summer at sea growth (FMC to FSM)
  alg_marked$FS.incr[i]<-sum(spacingavg[FMC:FSM])
  
  # First winter at sea growth (first circuli after FSM to M1)
  alg_marked$FW.incr[i]<-sum(spacingavg[(FSM+1):M1])
  
  # First winter minimum (FWM)
  # if there is only one marine annulus, stop searching at the end of the spacing vector
  # if there are two annuli, stop searching 10 circuli past the first marine annulus
  stopFWMsearch <- ifelse(is.na(homewater_human$M2[i]) == T, length(spacingavg), (M1 + 10))
  FWM<-(FSM-1)+which(spacingavg[FSM:stopFWMsearch]==min(spacingavg[FSM:stopFWMsearch]))[1]
  
  # First winter minimum (FWM) circulus number
  alg_marked$FWM.circ[i]<-FWM
  alg_marked$FWM[i]<-vect[FWM]
  
  # Post-smolt growth (FMC to M1)
  alg_marked$PS.incr[i]<-sum(spacingavg[FMC:M1])
  
  
  # First summer, number of circuli
  alg_marked$FS.circ[i]<-length(spacingavg[FMC:FSM])
  
  # First winter, number of circuli
  alg_marked$FW.circ[i]<-length(spacingavg[(FSM+1):M1])
  
  # Post-smolt, number of circuli
  alg_marked$PS.circ[i]<-length(spacingavg[FMC:M1])
  
  
  ####################################################################################
  
  # Depending on which growth markers are present, certain growth increments will be identifiable
  
  # is M2 present? if not, go to next row
  if(is.na(homewater_human$M2[i])){
    next
  }
  
  ####################################################################################
  
  # if M2 is present, calculate these growth markers, increments and circuli
  
  # M2 circuli number if present
  M2<-(which(homewater_human$M2[i]<=vect)[1])-1
  alg_marked$M2[i]<-M2
  
  # Second summer maximum intercirucli spacing between M1 and M2
  # (add 1 to M1 so as to not include M1 in second-year)
  # find the max value between the first and second marine annuli
  SSM<-M1+which(spacingavg[(M1+1):M2]==max(spacingavg[(M1+1):M2]))[1]
  
  
  # Second summer maximum (SSM) circulus number
  alg_marked$SSM.circ[i]<-SSM
  alg_marked$SSM[i]<-vect[SSM]
  
  # Second summer at sea growth (first circuli after M1 to SSM)
  alg_marked$SS.incr[i]<-sum(spacingavg[(M1+1):SSM])
  
  # Second winter at sea growth (first circuli after SSM to M2)
  alg_marked$SW.incr[i]<-sum(spacingavg[(SSM+1):M2])
  
  
  
  
  # Second winter minimum (SWM)
  SWM<-(SSM-1)+which(spacingavg[SSM:length(spacingavg)]==min(spacingavg[SSM:length(spacingavg)]))[1]
  
  # Second winter minimum (SWM) circulus number
  alg_marked$SWM.circ[i]<-SWM
  alg_marked$SWM[i]<-vect[SWM]
  
  
  
  
  
  # Year 2 growth (first circuli after M1 to M2)
  alg_marked$Y2.incr[i]<-sum(spacingavg[(M1+1):M2])
  
  # Second summer, number of circuli
  alg_marked$SS.circ[i]<-length(spacingavg[(M1+1):SSM])
  
  # Second winter, number of circuli
  alg_marked$SW.circ[i]<-length(spacingavg[(SSM+1):M2])
  
  # Year-two number of circuli
  alg_marked$Y2.circ[i]<-length(spacingavg[(M1+1):M2])
  
}

```

Create some "test" columns to validate calculations (human)

```{r}
human_marked_test<-human_marked %>% 
  mutate(testcol1 = round(total.incr - (smolt.incr + marine.incr),1)) %>% 
  mutate(testcol2 = total.circ - (smolt.circ + marine.circ)) %>% 
  mutate(testcol3 = round(PS.incr - (FS.incr + FW.incr),1)) %>% 
  mutate(testcol4 = Y2.circ - (SS.circ + SW.circ))

hist(human_marked_test$testcol1)
hist(human_marked_test$testcol2)
hist(human_marked_test$testcol3)
hist(human_marked_test$testcol4)
```

Create some "test" columns to validate calculations (algorithm)

```{r}
alg_marked_test<-alg_marked %>% 
  mutate(testcol1 = round(total.incr - (smolt.incr + marine.incr),1)) %>% 
  mutate(testcol2 = total.circ - (smolt.circ + marine.circ)) %>% 
  mutate(testcol3 = round(PS.incr - (FS.incr + FW.incr),1)) %>% 
  mutate(testcol4 = Y2.circ - (SS.circ + SW.circ))

hist(alg_marked_test$testcol1)
hist(alg_marked_test$testcol2)
hist(alg_marked_test$testcol3)
hist(alg_marked_test$testcol4)
```

Combine the human and algorithm data into one file, and write it to the processed data folder

```{r}
alg_vs_human_markers<-bind_rows(human_marked, alg_marked) %>% 
  drop_na() 

## write.csv(alg_vs_human_markers, here::here("processed_data","alg_vs_human_markers.csv"))

```















